\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual All-Loop Integrands

### 2023-08-31 notes/2loopNLSM.tex
Working on pathological def.

In the context of loop integrand construction, generalized unitarity
allows us to equate products of on-shell tree amplitudes, encoded via
a graph $\gamma$, to sums over compatible\footnote{A diagram $g$ is compatible with the
  cut diagram $\gamma$ if $g$ is isomorphic to $\gamma$ or to any of
  the additional factorization channels of $\gamma$. } diagram
numerators evaluated on the support of the on-shell conditions and
weighted by their uncut propagators
\begin{equation}
\label{eq:SchematicCutEquation}
\text{Cut}[\gamma]= \sum_{\substack{\text{states}\\\text{crossing } E(\gamma)}} \prod_{v \in V(\gamma)} \atree(v)
= \sum_{g \text{ compat } \gamma}
\frac{n[g] \vert_\text{cut}}{\prod \limits_{\text{uncut}} \ell^2}
 \, ,
\end{equation}
where $V(\gamma)$ are the vertices of $\gamma$ and $E(\gamma)$ its
edges.  The sum over states hides much of the complexity of cut
construction: for scalars the process is trivial while spinning states
greatly complicate matters (see
Refs.~\cite{Kosmopoulos:2020pcd,Edison:2022jln} for recent discussion
and detailed examples of evaluating state sums).  It is useful to
provide some examples of the two types of expressions appearing in
\cref{eq:SchematicCutEquation}.  First the so-called
iterated unitarity cut, which can be written in the product-of-trees
representation as
\begin{equation}
  \label{eq:example-cut-prod}
  \ExCutLabeled{1}{2}{3}{4} = \!\!\!\!\! \sum_{\substack{\text{states of }\\ \{\ell_1, \ell_2, \ell_3, \ell_4\}}} \!\!\!\!\!\!\!
  \atree(1,2,\ell_1,\ell_2) \atree(-\ell_1,-\ell_2,\ell_3,\ell_4) \atree(-\ell_3,-\ell_4,3,4) \,,
\end{equation}
or in reference to a collection of diagram numerators as
\begin{align}
  \label{eq:example-cut-sum}
  \ExCutLabeled{1}{2}{3}{4} &=
  \frac{\delta(\textcolor{cut-color}{\ell_1^2})
    \delta(\textcolor{cut-color}{\ell_2^2})
    \delta(\textcolor{cut-color}{\ell_3^2})
    \delta(\textcolor{cut-color}{\ell_4^2})}{(\textcolor{ucp-color}{\ell_5})^2
    (\textcolor{ucp-color}{\ell_6})^2
    (\textcolor{ucp-color}{\ell_7})^2}
  \ n\left[
    \IUCContrib{1}{4}{3}{2}
                              \right] \notag \\
  &\quad + 
  \frac{\delta(\textcolor{cut-color}{\ell_1^2})
    \delta(\textcolor{cut-color}{\ell_2^2})
    \delta(\textcolor{cut-color}{\ell_3^2})
    \delta(\textcolor{cut-color}{\ell_4^2})}
    {(\textcolor{ucp-color}{\ell_5})^2
    (\textcolor{ucp-color}{\ell_6})^2
    (\textcolor{ucp-color}{\ell_7'})^2}
  \ n\left[
    \IUCContrib{1}{3}{4}{2}
    \right] + 25\text{ others.}
\end{align}
When drawing cut diagrams, we will use gray blobs as in the left-hand
sides of \cref{eq:example-cut-prod,eq:example-cut-sum} and all drawn
legs are assumed to be on-shell.  When drawing numerator contributions
to cuts, we will not use blobs on vertices, and will use
\textcolor{cut-color}{dashed lines} bisecting edges to denote cut
lines and highlight the \textcolor{ucp-color}{uncut edges}.  From now
on we will leave the cut $\delta(\ell^2)$s implicit.

In color-charged theories, both sides of the cut can be further
decomposed according to the color algebra.  For theories charged under
the adjoint of $SU(N)$ via $f^{abc}$-dressed amplitudes

>>>>

In the context of loop integrand construction, generalized unitarity
allows us to equate products of on-shell tree amplitudes, encoded via
a graph $\gamma$, to sums over compatible\footnote{A diagram $g$ is compatible with the
  cut diagram $\gamma$ if $g$ is isomorphic to $\gamma$ or to any of
  the additional factorization channels of $\gamma$. } diagram
numerators evaluated on the support of the on-shell conditions and
weighted by their uncut propagators
\begin{equation}
\label{eq:SchematicCutEquation}
\text{Cut}[\gamma]= \sum_{\substack{\text{states}\\\text{crossing } E(\gamma)}} \prod_{v \in V(\gamma)} \atree(v)
= \sum_{g \text{ compat } \gamma}
\frac{n[g] \vert_\text{cut}}{\prod \limits_{\text{uncut}} \ell^2}
 \, ,
\end{equation}
where $V(\gamma)$ are the vertices of $\gamma$ and $E(\gamma)$ its
edges.  The sum over states hides much of the complexity of cut
construction: for scalars the process is trivial while spinning states
greatly complicate matters (see
Refs.~\cite{Kosmopoulos:2020pcd,Edison:2022jln} for recent discussion
and detailed examples of evaluating state sums).  It is useful to
provide some examples of the two types of expressions appearing in
\cref{eq:SchematicCutEquation}.  First the so-called
iterated unitarity cut, which can be written in the product-of-trees
representation as
\begin{equation}
  \label{eq:example-cut-prod}
  \ExCutLabeled{1}{2}{3}{4} = \!\!\!\!\! \sum_{\substack{\text{states of }\\ \{\ell_1, \ell_2, \ell_3, \ell_4\}}} \!\!\!\!\!\!\!
  \atree(1,2,\ell_1,\ell_2) \atree(-\ell_1,-\ell_2,\ell_3,\ell_4) \atree(-\ell_3,-\ell_4,3,4) \,,
\end{equation}
or in reference to a collection of diagram numerators as
\begin{align}
  \label{eq:example-cut-sum}
  \ExCutLabeled{1}{2}{3}{4} &=
  \frac{\delta(\textcolor{cut-color}{\ell_1^2})
    \delta(\textcolor{cut-color}{\ell_2^2})
    \delta(\textcolor{cut-color}{\ell_3^2})
    \delta(\textcolor{cut-color}{\ell_4^2})}{(\textcolor{ucp-color}{\ell_5})^2
    (\textcolor{ucp-color}{\ell_6})^2
    (\textcolor{ucp-color}{\ell_7})^2}
  \ n\left[
    \IUCContrib{1}{4}{3}{2}
                              \right] \notag \\
  &\quad + 
  \frac{\delta(\textcolor{cut-color}{\ell_1^2})
    \delta(\textcolor{cut-color}{\ell_2^2})
    \delta(\textcolor{cut-color}{\ell_3^2})
    \delta(\textcolor{cut-color}{\ell_4^2})}
    {(\textcolor{ucp-color}{\ell_5})^2
    (\textcolor{ucp-color}{\ell_6})^2
    (\textcolor{ucp-color}{\ell_7'})^2}
  \ n\left[
    \IUCContrib{1}{3}{4}{2}
    \right] + 25\text{ others.}
\end{align}
When drawing cut diagrams, we will use gray blobs as in the left-hand
sides of \cref{eq:example-cut-prod,eq:example-cut-sum} and all drawn
legs are assumed to be on-shell.  When drawing numerator contributions
to cuts, we will not use blobs on vertices, and will use
\textcolor{cut-color}{dashed lines} bisecting edges to denote cut
lines and use colored edges to highlight the
\textcolor{ucp-color}{uncut propagators}.  From now on we will leave
the cut $\delta(\ell^2)$s implicit.

In color-charged theories, both sides of the cut can be further
decomposed according to the color algebra.  For theories charged under
the adjoint of $SU(N)$ via $f^{abc}$-dressed amplitudes, it is
convenient to project onto a specific basis of the Del
Duca-Dixon-Maltoni (DDM) color basis \cite{DixonMaltoni} by only
inserting color-ordered amplitudes in the product of trees, and
restricting the set of compatible diagrams to those whose color
factors reduce to the appropriate DDM element when summing over
\emph{only the uncut color contractions}.  In this situation, the
color becomes an overall factor on the entire equality and thus can be
ignored.  Cuts that are organized in this manner are known as
``color-ordered'' cuts, in the same sense as color-ordered tree
amplitudes.  Since the DDM basis (and dual Kleiss-Kuijf amplitude
basis \cite{Kleiss:1988ne}) have $(n-2)!$ elements, one generally
needs to evaluate that many separate color-orderings of the same cut.
As an example, consider the ``kissing triangle'' cut topology, one of
the factorization channels of the iterated unitarity cut, which has
two topologically distinct color-ordered expansions
\begin{equation}
  \label{eq:kt-color-ord}
  \KissingTriangles =
  \begin{cases}
     s\text{-}t \text{ channel}
    &
      \frac{1}{\textcolor{ucp-color}{s_{12}}}
      n \left[\scalebox{0.65}{\KTBTContrib{1}{4}{3}{2}}
      \right]
      +
      \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
      n \left[
      \scalebox{0.65}{\KTDBContrib{1}{4}{3}{2}}
      \right]\\
    t\text{-}u \text{ channel}
    & 
      \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
      n \left[
      \scalebox{0.65}{\KTDBContrib{1}{4}{3}{2}}
      \right]
      +
      \frac{1}{\textcolor{ucp-color}{\ell_2^2}} \
      n\left[
      \scalebox{0.65}{\KTDBContrib{1}{3}{4}{2}}
      \right]
  \end{cases}
\end{equation}
which serve as a complete basis for the full cut.

Even beginning at one loop, we will encounter types of cuts that are
difficult to properly define due to hidden non-physical singularities.
The quintessential example of such a ``pathological'' cut is the 3-1
two-particle cut at one loop which hides a bubble-on-external-leg
singularity
\begin{equation}
  \begin{tikzpicture}
\begin{feynman}
\vertex (a1) at (-.8, -0.6) {};
\vertex (a2) at (-.8, 0.6) {};
\vertex (a3) at (-1, 0) {};
\vertex [dot, scale=2](mid1) at (0,0){};
\vertex [dot, scale=1.5,hgrey0](mid2) at (0,0){};
\vertex [dot, scale=2](mid3) at (1,0){};
\vertex [dot, scale=1.5,hgrey0](mid4) at (1,0){};
\vertex (a4) at (1.8, 0) {};
\diagram{
(mid1) --[ultra thick,](a1),
(mid1) --[ultra thick,](a2),
(mid1) --[ultra thick,](a3),
%(mid1) --[ultra thick,](mid3),
(mid1) --[ultra thick,out=60,in=120,min distance=0.4cm](mid3),
(mid1) --[ultra thick,out=-60,in=-120,min distance=0.4cm](mid3),
(mid3) --[ultra thick](a4),
};
\end{feynman}
\end{tikzpicture}
\ni
\frac{\delta(\textcolor{cut-color}{k_1^2})}{\textcolor{ucp-color}{k_1^2}}
n \left[
  \begin{tikzpicture}
    \ver{e1}{-1,0};
    \ver{e2}{-0.6,0.4};
    \ver{e3}{-0.3,0.4};
    \ver{i1}{-0.6,0};
    \ver{i2}{-0.3,0};
    \ver{i3}{0,0};
    \ver{i4}{0.8,0};
    \ver{co}{$(i3)!0.5!(i4)$};
    \ver{e4}{1.2,0};
    \draw[ultra thick] (e1)--(i1)--(i2)--(i3) to [out=80,in=100,min distance=0.4cm](i4)--(e4);
    \draw[ultra thick,ucp-color](i1)--(i2)--(i3);
    \draw[ultra thick,cut-color](i4)--(e4);
    \draw[ultra thick] (i3) to [out=-80,in=-100,min distance=0.4cm](i4);
    \draw[ultra thick] (e2) -- (i1);
    \draw[ultra thick] (e3)--(i2);
    \draw[ultra thick,dashed,cut-color] (co)--++(90:0.5);
    \draw[ultra thick,dashed,cut-color] (co)--++(-90:0.5);
  \end{tikzpicture}
  \right]
\end{equation}
